High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-04-27 notebooks/mnist.qmd
mnist benchmark set up

# Data:
n_obs = 1000
counterfactual_data = load_mnist(n_obs)
counterfactual_data.X = pre_process.(counterfactual_data.X)
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
x_factual = reshape(pre_process(x_factual, noise=0.0f0), input_dim, 1)
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
n_digits = Int(sqrt(input_dim))
output_dim = length(unique(labels))
```

@@

```{julia}
# JEM parameters:
𝒟x = Uniform(0,1)
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=10,
)
α = [1.0,1.0,1e-1]      # penalty strengths
```

@@

```{julia}
# Plot generated samples:
if _regen 
    for (mod_name, mod) in model_dict
        if ECCCo._has_sampler(mod)
            sampler = ECCCo._get_sampler(mod)
        else
            K = length(counterfactual_data.y_levels)
            input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
            𝒟x = Uniform(extrema(counterfactual_data.X)...)
            𝒟y = Categorical(ones(K) ./ K)
            sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size)
        end
        opt = ImproperSGLD()
        f(x) = logits(mod, x)

        n_iter = 200
        _w = 1500
        plts = []
        neach = 10
        for i in 1:10
            x = sampler(f, opt; niter=n_iter, n_samples=neach, y=i)
            plts_i = []
            for j in 1:size(x, 2)
                xj = x[:,j]
                xj = reshape(xj, (n_digits, n_digits))
                plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
            end
            plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
            plts = [plts..., plt]
        end
        plt = Plots.plot(plts..., size=(_w,_w), layout=(10,1), plot_title=mod_name)
        savefig(plt, joinpath(output_images_path, "mnist_generated_$(mod_name).png"))
        display(plt)
    end
end
```

@@

```{julia}
function _plot_eccco_mnist(
    x::Union{AbstractArray, Int}=x_factual, target::Int=target;
    λ=[0.1,0.1,0.1],
    temp=0.1,η=0.01,
    plt_order = ["MLP", "MLP Ensemble", "JEM", "JEM Ensemble"],
    opt = Flux.Optimise.Adam(η),
    rng::Union{Int,AbstractRNG}=Random.GLOBAL_RNG,
)

@@

```{julia}
# Measures:
measures = [
    CounterfactualExplanations.distance,
    ECCCo.distance_from_energy,
    ECCCo.distance_from_targets,
    CounterfactualExplanations.validity,
    CounterfactualExplanations.redudancy,
]
```

>>>>

# Data:
n_obs = 10000
counterfactual_data = load_mnist(n_obs)
counterfactual_data.X = pre_process.(counterfactual_data.X)
counterfactual_data.generative_model = vae
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
x_factual = reshape(pre_process(x_factual, noise=0.0f0), input_dim, 1)
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
n_digits = Int(sqrt(input_dim))
output_dim = length(unique(labels))
```

@@

```{julia}
# JEM parameters:
𝒟x = Uniform(0,1)
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=10,
)
α = [1.0,1.0,1e-2]      # penalty strengths
```

@@

```{julia}
# Plot generated samples:
n_regen = 150
if _regen 
    for (mod_name, mod) in model_dict
        if ECCCo._has_sampler(mod)
            sampler = ECCCo._get_sampler(mod)
        else
            K = length(counterfactual_data.y_levels)
            input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
            𝒟x = Uniform(extrema(counterfactual_data.X)...)
            𝒟y = Categorical(ones(K) ./ K)
            sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size)
        end
        opt = ImproperSGLD()
        f(x) = logits(mod, x)

        _w = 1500
        plts = []
        neach = 10
        for i in 1:10
            x = sampler(f, opt; niter=n_regen, n_samples=neach, y=i)
            plts_i = []
            for j in 1:size(x, 2)
                xj = x[:,j]
                xj = reshape(xj, (n_digits, n_digits))
                plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
            end
            plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
            plts = [plts..., plt]
        end
        plt = Plots.plot(plts..., size=(_w,_w), layout=(10,1), plot_title=mod_name)
        savefig(plt, joinpath(output_images_path, "mnist_generated_$(mod_name).png"))
        display(plt)
    end
end
```

@@

```{julia}
function _plot_eccco_mnist(
    x::Union{AbstractArray, Int}=x_factual, target::Int=target;
    λ=[0.5,0.1,0.5],
    temp=0.1,η=0.01,
    plt_order = ["MLP", "MLP Ensemble", "JEM", "JEM Ensemble"],
    opt = Flux.Optimise.Adam(η),
    rng::Union{Int,AbstractRNG}=Random.GLOBAL_RNG,
)

@@

```{julia}
# Measures:
measures = [
    CounterfactualExplanations.distance,
    ECCCo.distance_from_energy,
    ECCCo.distance_from_targets,
    CounterfactualExplanations.Evaluation.validity,
    CounterfactualExplanations.Evaluation.redundancy,
]

bmk = benchmark(
    counterfactual_data; 
    models=model_dict, 
    generators=generator_dict, 
    measure=measures,
    suppress_training=true, dataname="MNIST",
    n_individuals=100,
    initialization=:identity,
)

Serialization.serialize(joinpath(output_path, "mnist_benchmark.jls"), bmk)
```

```{julia}
@chain bmk() begin
    @group_by(dataname, generator, model, variable)
    @summarize(mean=mean(value),sd=std(value))
    @ungroup
    @filter(variable == "distance_from_targets")
end
```
